---
title: "SoloLearn Data Science"
author: "Ken Harmon"
date: "`r format(Sys.time(), '%Y %B %d')`"
output:
  html_document:
    keep_md: yes
    code_folding: hide
    fig_height: 6
    fig_width: 12
    fig_align: center
  pdf_document: default
editor_options:
  chunk_output_type: console
---
 
https://www.sololearn.com/Play/machine-learning

# {.tabset .tabset-fade}

```{r, echo=FALSE}
library(reticulate)
knitr::opts_chunk$set(echo = TRUE, message = FALSE, warning = FALSE)
knitr::knit_engines$set(python = reticulate::eng_python)
reticulate::repl_python()

```

```{r}
#py_install("graphviz")
#py_install("sqlalchemy")
```

https://www.sololearn.com/Play/data-science

Welcome to Data Science

Congratulations on taking a big step toward becoming a data scientist!
In addition to working through this course, be sure to take advantage of all of the learning support available to you on SoloLearn, including the daily tips, Try it Yourself practices, code coach challenges, code playground, and engagement with our amazing learner community. We love to hear from you, so please leave comments and feedback as you learn with us.

What is Data Science?

There are many use cases in business for data science including finding a better housing price prediction algorithm for Zillow, finding key attributes associated with wine quality, and building a recommendation system to increase the click-through-rate for Amazon.

Extracting insights from seemingly random data, data science normally involves collecting data, cleaning data, performing exploratory data analysis, building and evaluating machine learning models, and communicating insights to stakeholders.

Why Python?

In this Introduction to Data Science course we’re learning data science with Python. As a general-purpose programming language, Python is now the most popular programming language in data science. It’s easy to use, has great community support, and integrates well with other frameworks (e.g., web applications) in an engineering environment.

This course focuses on exploratory data analysis with three fundamental Python libraries: numpy, pandas and matplotlib. The machine learning library scikit-learn will be covered as well.

In the later modules, we will be predicting home values using linear regression, identifying classes of iris with classification algorithms, and finding clusters within wines, just a few examples of what we can do in data science.
In data science, there are other popular programming languages, such as R, which has an edge in statistical modeling.

Numerical Data

Datasets come from a wide range of sources and formats: it could be collections of numerical measurements, text corpus, images, audio clips, or basically anything. No matter the format, the first step in data science is to transform it into arrays of numbers.

We collected 45 U.S. president heights in centimeters in chronological order and stored them in a list, a built-in data type in python.


```{python}
heights = [189, 170, 189, 163, 183, 171, 185, 168, 173, 183, 173, 173, 175, 178, 183, 193, 178, 173, 174, 183, 183, 180, 168, 180, 170, 178, 182, 180, 183, 178, 182, 188, 175, 179, 183, 193, 182, 183, 177, 185, 188, 188, 182, 185, 191]
```

In this example, George Washington was the first president, and his height was 189 cm.

If we wanted to know how many presidents are taller than 188cm, we could iterate through the list, compare each element against 188, and increase the count by 1 as the criteria is met.

```{python}
cnt = 0
for height in heights:
  if height > 188:
    cnt +=1
print(cnt)
```

This shows that there are five presidents who are taller than 188 cm.

Introduction to Numpy

Numpy (short for Numerical Python) allows us to find the answer to how many presidents are taller than 188cm with ease. Below we show how to use the library and start with the basic object in numpy.

```{python}
import numpy as np
heights_arr = np.array(heights)
print((heights_arr > 188).sum())
```

The import statement allows us to access the functions and modules inside the numpy library. The library will be used frequently, so by convention numpy is imported under a shorter name, np. The second line is to convert the list into a numpy array object, via np.array(), that tools provided in numpy can work with. The last line provides a simple and natural solution, enabled by numpy, to the original question.

As our datasets grow larger and more complicated, numpy allows us the use of a more efficient and for-loop-free method to manipulate and analyze our data. Our dataset example in this module will include the US Presidents' height, age and party.

Size and Shape

An array class in Numpy is called an ndarray or n-dimensional array. We can use this to count the number of presidents in heights_arr, use attribute numpy.ndarray.size:

```{python}
heights_arr.size
```

Note that once an array is created in numpy, its size cannot be changed.

Size tells us how big the array is, shape tells us the dimension. To get current shape of an array use attribute shape:

```{python}
heights_arr.shape
```

The output is a tuple, recall that the built-in data type tuple is immutable whereas a list is mutable, containing a single value, indicating that there is only one dimension, i.e., axis 0. Along axis 0, there are 45 elements (one for each president) Here, heights_arr is a 1d array.
Attribute size in numpy is similar to the built-in method len in python that is used to compute the length of iterable python objects like str, list, dict, etc.

Reshape

Other data we have collected includes the ages of the presidents:


```{python}
ages = [57, 61, 57, 57, 58, 57, 61, 54, 68, 51, 49, 64, 50, 48, 65, 52, 56, 46, 54, 49, 51, 47, 55, 55, 54, 42, 51, 56, 55, 51, 54, 51, 60, 62, 43, 55, 56, 61, 52, 69, 64, 46, 54, 47, 70]
```

Since both heights and ages are all about the same presidents, we can combine them:

```{python}
heights_and_ages = heights + ages 
# convert a list to a numpy array
heights_and_ages_arr = np.array(heights_and_ages)
heights_and_ages_arr.shape
```

This produces one long array. It would be clearer if we could align height and age for each president and reorganize the data into a 2 by 45 matrix where the first row contains all heights and the second row contains ages. To achieve this, a new array can be created by calling numpy.ndarray.reshape with new dimensions specified in a tuple:

```{python}
heights_and_ages_arr.reshape((2,45))
```

The reshaped array is now a 2darray, yet note that the original array is not changed. We can reshape an array in multiple ways, as long as the size of the reshaped array matches that of the original.

```{python}
newP = heights_and_ages_arr.reshape((2,45))
newP.shape
newP
```

Data Type

Another characteristic about numpy array is that it is homogeneous, meaning each element must be of the same data type.

For example, in heights_arr, we recorded all heights in whole numbers; thus each element is stored as an integer in the array. To check the data type, use numpy.ndarray.dtype

```{python}
heights_arr.dtype
```

If we mixed a float number in, say, the first element is 189.0 instead of 189

```{python}
heights_float = [189.0, 170, 189, 163, 183, 171, 185, 168, 173, 183, 173, 173, 175, 178, 183, 193, 178, 173, 174, 183, 183, 180, 168, 180, 170, 178, 182, 180, 183, 178, 182, 188, 175, 179, 183, 193, 182, 183, 177, 185, 188, 188, 182, 185, 191]
```

Then after converting the list into an array, we’d see all other numbers are coerced into floats:

```{python}
heights_float_arr = np.array(heights_float)
heights_float_arr
heights_float_arr.dtype
```




















